---
title: "Hierarchical Clustering"
output:
  html_document:
    keep_md: yes
---

```{r, message=FALSE}
library(gplots)
library(pheatmap)
library(dplyr)

source('helpers.R', local=TRUE)
```

## Load dataset [1, 2]
```{r, message=FALSE, cache=TRUE}
gds4198 <- get_gds4198()
```

## Perform hierarchical clustering
```{r}
gds4198_samples_dm <- dist(t(gds4198$data))
gds4198_samples_hc <- hclust(gds4198_samples_dm)
```

## Dendrogram

```{r, fig.width=10}
plot(dendrapply(
    as.dendrogram(gds4198_samples_hc),
    color_by_subtype,
    colnames(gds4198$data), gds4198$subtypes)
)
title('Dendrogram GDS4198, complete linkage')
```

## Heatmap

```{r, fig.width=10, fig.height=10}
expr_df <- data.frame(gene = gds4198$genes, gds4198$data) %>%
    group_by(gene) %>% summarise_each(funs(mean))
expr_df <- expr_df %>% filter(gene %in% top_accross_conditions(expr_df))
expr_matrix <- data.matrix(expr_df[, -1])
rownames(expr_matrix) <- expr_df$gene

gplots::heatmap.2(
    expr_matrix, trace='none', density.info='none',
    col=colorRampPalette(c('#f0f0f0', '#636363'))(70),
    ColSideColors=c('#66c2a5', '#fc8d62', '#8da0cb')[gds4198$subtypes]
)
```

## Pretty heatmap

```{r, fig.width=10, fig.height=10}
ann <- data.frame(subtype=gds4198$subtypes)
rownames(ann) <- colnames(gds4198$data)
ann_colors <- list(subtype=setNames(
    c('#66c2a5', '#fc8d62', '#8da0cb'),
    levels(gds4198$subtypes)
))

pheatmap::pheatmap(
    expr_matrix, annotation=ann,
    annotation_colors=ann_colors
)
```


---------------------------------------


1. Wu, Y. et al. Comprehensive genomic meta-analysis identifies intra-tumoural stroma as a predictor of survival in patients with gastric cancer. Gut 62, 1100–1111 (2013).
2. Chia, N.-Y. et al. Regulatory crosstalk between lineage-survival oncogenes KLF5, GATA4 and GATA6 cooperatively promotes gastric cancer development. Gut (2014). doi:10.1136/gutjnl-2013-306596
