---
title: "Principal component analysis"
output:
  html_document:
    keep_md: yes
---

```{r, message=FALSE}
library(ggplot2)
library(gridExtra)

# For get_gds4198 and plot_pc
source('helpers.R', local=TRUE)
```

## Load dataset [1, 2]

```{r load_gds4198, message=FALSE, cache=TRUE}
gds4198 <- get_gds4198()
```

Dataset contains `r ncol(gds4198$data)` samples from 3 subtypes (`r paste(levels(gds4198$subtypes), collapse=', ')`) of the primary gastric tumors.


## Apply principal component analysis

R `stats` library provides two functions performing principal component analysis: `princomp` and `prcomp`. Since `princomp` requires more observations than variable we'll use `prcomp`.

```{r}
gds4198_pc <- prcomp(t(gds4198$data))
```

Check percent of variance explained by first 3 prinicpal components.

```{r}
explained_variance <- (gds4198_pc$sdev) ^ 2 / sum(gds4198_pc$sdev ^ 2) * 100

round(sum(explained_variance[1:3]), 3)
```

Extract cooridnates of observations in the principal component space and plot first two.

```{r, fig.height=10, fig.width=10}
gds4198_pc_x <- as.data.frame(gds4198_pc$x)
gds4198_pc_x$subtype <- gds4198$subtypes

plot_pc(x=gds4198_pc_x, main='GDS4198 - first three principal components')
```

1. Wu, Y. et al. Comprehensive genomic meta-analysis identifies intra-tumoural stroma as a predictor of survival in patients with gastric cancer. Gut 62, 1100–1111 (2013).
2. Chia, N.-Y. et al. Regulatory crosstalk between lineage-survival oncogenes KLF5, GATA4 and GATA6 cooperatively promotes gastric cancer development. Gut (2014). doi:10.1136/gutjnl-2013-306596